# Introduction

Welcome to the pOmics Handbook. 



## Setup R and RStudio

First, we need to install the programming language R and then RStudio, the most commonly used code editor for R. Download and install the most recent version of R for your platform from the [R Project website](https://www.r-project.org/). 

After successful installation of R, download RStudio. The free [RStudio Desktop](https://www.rstudio.com/products/rstudio/) Open Source Edition will suffice for all your needs.

If you are having trouble, you can find a more comprehensive explanations [here](https://moderndive.netlify.app/1-getting-started.html).


## R Basics

While there are many introductions to R out in the internet like [R for Data Science](https://r4ds.had.co.nz/introduction.html), the [swirl](https://swirlstats.com/) package, some introductions from [Modern Dive](https://moderndive.netlify.app/index.html), and other resources you can find on the official website of the [bookdown](https://bookdown.org/) package, this chapter might be filled with tips that help users getting start coding in R.

If none of the above material appeals to you, try [this](https://letmegooglethat.com/?q=introduction+to+r).


## Install R packages







## Install R Bioconductor  packages

Install devtools using `install.packages("devtools")`.

Install Bioconductor (https://bioconductor.org/) via 
```{r bioc, eval=FALSE, message=F}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()
```



Install the tidyverse (https://www.tidyverse.org/) via
```{r install tidyR, eval=FALSE}
install.packages("tidyverse")
```
