# Getting started with R and RStudio


## Install the R language

Download the most recent release of R for your platform from https://cloud.r-project.org/. Install it as any other program. It is important to install R before RStudio.


## Install RStudio

Download the desktop version of RStudio from https://www.rstudio.com/products/rstudio/. This is the program in which you will interact with the R language and conduct all your analyses.


## Install others

To install packages from GitHub, you need RTools. Install RTools from https://cran.r-project.org/bin/windows/Rtools/.




## Install R packages

"R packages are collections of functions and data sets developed by the community. They increase the power of R by improving existing base R functionalities, or by adding new ones."^[https://www.datacamp.com/community/tutorials/r-packages-guide] 


(Available CRAN Packages By Name)[https://cran.r-project.org/web/packages/available_packages_by_name.html]



## Install R Bioconductor  packages

Install devtools using `install.packages("devtools")`.

Install Bioconductor (https://bioconductor.org/) via 
```{r bioc, eval=FALSE, message=F}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install()
```



Install the tidyverse (https://www.tidyverse.org/) via
```{r install tidyR, eval=FALSE}
install.packages("tidyverse")
```
