# Getting started with R and RStudio


## Install the R language

Download the most recent release of R for your platform from https://cloud.r-project.org/. Install it as any other program. It is important to install R before RStudio.


## Install RStudio

Download the desktop version of RStudio from https://www.rstudio.com/products/rstudio/. This is the program in which you will interact with the R language and conduct all your analyses.



## Install R packages

"R packages are collections of functions and data sets developed by the community. They increase the power of R by improving existing base R functionalities, or by adding new ones."^[https://www.datacamp.com/community/tutorials/r-packages-guide] 


[Available CRAN Packages By Name](https://cran.r-project.org/web/packages/available_packages_by_name.html)



Install the tidyverse (https://www.tidyverse.org/) via
```{r install tidyR, eval=FALSE}
install.packages("tidyverse")
```

Next, we will download two more packages, which will help us install even more R packages from sources other than CRAN.
```{r install more, eval=FALSE}
install.packages(c("devtools", 
                   "BiocManager"))
```



## Install R from Bioconductor


Install Bioconductor (https://bioconductor.org/) via 
```{r bioc, eval=FALSE, message=F}
BiocManager::install(c("clusterProfiler",
                       "org.Hs.eg.db"
                       "UniProt.ws"))
```







## Install R packages from GitHub and others

Another important source of R packages is GitHub. GitHub is not just the place where most R packages are being developed before they are put to repositories such as CRAN or Bioconductor, many other packages including the `pOmics` packages can be installed from it. :)

Let's do it:
```{r install pOmics, eval=FALSE}
devtools::install_github("nicohuttmann/pOmics")
```




Last but not least, there are some other sources for R packages which will be exemplified by the `disgenet2r` package to query the [DisGeNET](https://www.disgenet.org/) database for gene disease associations.
```{r install disgenet2r, eval=FALSE}
devtools::install_bitbucket("ibi_group/disgenet2r")
```


